Universal Understanding of the Dielectric Nature of Two-Dimensional Materials

### 2018-11-15 SI.tex
Move all theoretical part to SI

\section{More theoretical analysis}
\label{eq:MG-effect-1}

\subsection{Detailed derivations for Eq. 5a and 5b}
\label{sec:theory-1}

Here we will first show how Eq. 5a and 5b in the main text are
derived. Taking the limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$ at 0 K can be unified by the
same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$.

For the in-plane component, $\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$
is independent of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes the independent of $\psi^{\perp}$, due to the
orthogonality. Present the summation of k-space into integral, and further
combining with the Bloch wave representation and the k-p
theory\cite{Jiang_2017_Eg_Eb}, we arrive at Eq. 5a.

For Eq. 5b, treating the in-plane wavefunctions as plane wave with
form $\psi^{\parallel}(\rho) \propto e^{i \mathbf{k \rho}}$, the
matrix element of
$<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>$,
when $\mathbf{q}=(0, 0, q_{z})$, becomes\cite{Hybertsen_1987}:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
  <\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>
  &= \frac{1}{A} \int dx \int dy
  e^{i(\mathbf{-k \rho} - \mathbf{q \rho} + \mathbf{(k+q) \rho})}
  \int (\psi^{\perp})^{*}_{\mathrm{v}}(\mathbf{k})e^{-iq_{z}z}\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})\\
  &= <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})>
  \end{aligned}
\end{equation}
Note that the states perpendicular are bound, the integral is
meaningful only when $\mathbf{k=k+q}$ \cite{davies_physics_1997}. By
performing the Taylor expansion of
$e^{-i\mathbf{qr}} \approx 1 - i\mathbf{qr}$, we get:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
    <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
    &\approx <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})> -
    iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>\\
    &= -iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
   \end{aligned}
\end{equation}
plug this into Eq. \ref{eq:alpha-RPA} and express the summation over
$k_{x}$ and $k_{y}$ in a continuous form within the Brillouin Zone, we
arrive at Eq. 5b in main text.

Although the exact form for $\psi^{\perp}$ depends on the exact
distribution of $V(z)$, without loss of generality we can assume the
electrons are confined in a potential well of width $\delta$, which is
the typical treatment for semiconductor QWs
\cite{ihn_semiconductor_2009}. The allowed bound states inside the
confined region generally has wave vector
$k_{z} \propto n \pi / \delta$. With the total energy
$E_{n}(\mathbf{k}) = {\displaystyle \frac{\hbar^{2} (k_{x}^{2} +
    k_{y}^{2})}{2 m^{\parallel}} + \frac{\hbar^{2} n^{2} \pi^{2}}{2
    m^{\perp} \delta^{2}}}$, where $m^{\parallel}$ and $m^{\perp}$ are
the effective masses parallel and perpendicular to the 2D
plane. Therefore the denominator of Eq. 10 becomes independent of
$\mathbf{k}$, that
$E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k}) =
(n_{\mathrm{c}}^{2} - n_{\mathrm{v}}^{2}) {\displaystyle
  \frac{\hbar^{2} \pi^{2}}{2 m^{\perp} \delta^{2}}}$. On the other
hand, the numerator
$<\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>$
is proportional to the confinement length $\delta$ (can be seen using
particle-in-box solution\cite{davies_physics_1997}). In combination,
the individual terms of the summation in the right hand of Eq. 5b is
independent of neither $E_{\mathrm{g}}$ nor \textbf{k}, proving that
$\alpha_{\mathrm{2D}}^{\perp}$ is independent of the band gap.

Finally we will show the $\alpha_{\mathrm{2D}}^{\perp}$ is related to the
confinement length $\delta$, or equivalent the thickness of the 2D
material. For simplicity we assume $\psi^{\parallel}$ is the solution
of a particle confined inside an infinite potential well of width
$\delta$ with the form $\psi \propto \sin(\frac{n\pi}{\delta})$. At 0
K, the only non-zero contribution for the transition between band
$n_{\mathrm{v}}$ and $n_{\mathrm{c}}$ corresponds with in-plane wave
vector $\mathbf{k}_{\rho}$ within the range of $k_{\mathrm{F}}^{\mathrm{v}}$ to
$k_{\mathrm{F}}^{\mathrm{c}}$, the Fermi vectors of each band. Therefore we have:
\begin{equation}
  \begin{aligned}
    \alpha_{\mathrm{2D}}^{\perp} &\propto \delta^{4} m^{\perp} \int
    d^{2}\mathbf{k}_{\rho}
    \frac{f(\psi_{\mathrm{v}}(\mathbf{k}_{\rho}))
      -f(\psi_{\mathrm{v}}(\mathbf{k}_{\rho})) }{n_{\mathrm{c}}^{2} -
      n_{\mathrm{v}}^{2}}\\
    &= \delta^{4} m^{\perp} \sum_{\mathrm{c, v}}\pi \frac{
      (k_{\mathrm{F}}^{\mathrm{c}})^{2} -
        (k_{\mathrm{F}}^{\mathrm{v}})^{2}}{n_{\mathrm{c}}^{2} -
        n_{\mathrm{v}}^{2}}
  \end{aligned}
\end{equation}
Since the Fermi wave vector for band \textit{i} follows
${\displaystyle
  \frac{\hbar^{2}(k_{\mathrm{F}}^{i})^{2}}{2m^{\parallel}}} +
{\displaystyle \frac{\hbar^{2} n_{i}^{2} \pi^{2}}{2m^{\perp}
    \delta^{2}}} = E_{\mathrm{F}}$, we have
$ (k_{\mathrm{F}}^{\mathrm{c}})^{2} -
(k_{\mathrm{F}}^{\mathrm{v}})^{2} \propto (n_{\mathrm{c}}^{2} -
n_{\mathrm{v}}^{2}) {\displaystyle
  \frac{m^{\parallel}}{m^{\perp}}}$. Note that in the particle-in-box
model, the energy levels for a sub-nm QW will be at the order of 10
eV, which allows only few states inside the QW. Combine with the fact
that the strongest transition comes from the lowest bands
\cite{davies_physics_1997}, we can finally show:
\begin{equation}
  \label{eq:alpha-perp-L}
  \alpha_{\mathrm{2D}}^{\perp} \propto \delta^{2}
\end{equation}
Apart from the power of $\delta$, this explains our observation
$\alpha_{\mathrm{2D}}^{\perp}$ is related with the thickness of a 2D material. We
note the power law of $\delta^{2}$ instead of $\delta$ is
inevitable. The atomic polarizabilities of free particles inside a 1D
infinite QW, finite QW or harmonic QW all show dependency with
$\delta^{4}$, where $\delta$ is the length of confinement or effective
harmonic length of the QW \cite{Fowler_1984,Maize_2011}. Since the 2D
polarizability is the molecular polarizability divided by the area,
such approach will always yield $\alpha_{\mathrm{2D}}^{\perp} \propto \delta^{2}$
regardless of the Hamiltonian used. To accurately model the
$\alpha_{\mathrm{2D}}^{\perp}$ we need a better model for the confinement in
z-direction than simple square potential well, which may be
tedious. Nevertheless the simple model presented here still captures
the bandgap-indepent and thickness-related feature of
$\alpha_{\mathrm{2D}}^{\perp}$, supporting our findings in the main text.

The dependency of $\alpha_{\mathrm{2D}}^{\perp}$ on the thickness of a 2D material,
can also be regarded using fundamental electrostatic model. Consider
the smallest repeating unit of the 2D material with xy-plane area $A$,
under small perturbation field $E$ along the z-direction.  Note that
the surface bound charge $\sigma_{\mathrm{b}}=n e /A$, where $n$ is
the number of unit charges contributes to the bound charges, comes
only from the dipoles of the outer-most atoms, since the induced
charges from inner atoms are cancel out (see Figure \ref{fig:classic-model}).
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7\linewidth]{img/SI-classic.pdf}
  \caption{Fundamental electrostatic model for the
    thickness-dependency of $\alpha_{\mathrm{2D}}^{\perp}$, using 2H-MoS2 as an
    example. Left: induced dipoles from individual atoms along the
    z-direction. The positive and negative induced charges from inner
    atoms cancel out. Right: simplied model for the thickness
    dependency of $\alpha_{\mathrm{2D}}^{\perp}$, where the surface dipole density
    $\boldsymbol{\mu}$ comes only from the outer-most atoms.}
  \label{fig:classic-model}
\end{figure}
From the definition of
$\alpha_{\mathrm{2D}}^{\perp}$, we have:
\begin{equation}
  \label{eq:alpha-classic}
  \alpha_{\mathrm{2D}}^{\perp} = \frac{\boldsymbol{u}_{z}}{E_{\mathrm{loc}} A}
  = \frac{(d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j}) \sigma_{\mathrm{b}}}{E}
\end{equation}
where $r_{\mathrm{cov}}^{i}$ and $r_{\mathrm{cov}}^{j}$ are the
covalent radii of the outer-most atoms, the characteristic length of
the dipole extension in z-direction, and $d_{\mathrm{max}}$ is the
z-distance between the nuclei of such atoms.  The field $E$
counterbalances the field from the surface bound charges and equals
$E = \sigma_{\mathrm{b}}/\varepsilon_{0}$. Therefore we have:
\begin{equation}
  \label{eq:alpha-classic-2}
  \alpha_{\mathrm{2D}}^{\perp} = (d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j})\varepsilon_{0}
                = \delta_{\mathrm{cov}} \varepsilon_{0}
\end{equation}
which explains the linear relation seen in Figure 3c of main text. We
can see that such simple model nicely captures the thickness feature
of $\alpha_{\mathrm{2D}}^{\perp}$, and reproduces the right coefficient between
$\delta_{\mathrm{cov}}$ and $\alpha_{\mathrm{2D}}^{\perp}$. Finally, the schematic
drawing of the polarizability ellipsoid for a 2D material, is drawn in
Figure \ref{fig:ellipsoid}.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{img/SI-ellipsoid.pdf}
  \caption{Polarizability ellipsoid of a 2D material, the long axis
    equals $\alpha_{\mathrm{2D}}^{\parallel}/2\varepsilon_{0}$ and the short axis
    equals  $\alpha_{\mathrm{2D}}^{\perp}/2\varepsilon_{0}$}
  \label{fig:ellipsoid}
\end{figure}

\subsection{Detailed discussion for main text Eq. 9}
\label{sec:theory-2}

>>>>

\section{Theoretical Analysis}
\label{sec:theory}
\subsection{Polarizability-Based Theoretical Model}
\label{ssec:theory-1}
In this section we show in detail the polarizability-based theoretical
framework that leads to the 2D Moss-like relations proposed in the
main text.  Due to its highly anisotropic nature, the wavefunction of
an isolated 2D material $\psi(\mathbf{r})$ can be separated into the
in- and out-of-plane components, similar to the treatment of quantum
wells (QW),\cite{davies_physics_1997} such that
$\psi^{\parallel}(\boldsymbol{\rho})$ and $\psi^{\perp}(z)$, such that
$\psi=\psi^{\parallel}\psi^{\perp}$, where $\boldsymbol{\rho}=(x, y)$
is the in-plane coordinate. Using the Bloch theorem, the periodic
$\psi^{\parallel}(\boldsymbol{\rho})$ can be further expressed as
$\psi^{\parallel}(\boldsymbol{\rho})=e^{i\mathbf{k} \cdot
  \boldsymbol{\rho}}u(\boldsymbol{\rho})$, where $\mathbf{k}$ is the
in-plane wave vector and $u(\boldsymbol{\rho})$ is periodic function
in the xy-plane. According to the random phase approximation (RPA)
theory\cite{Adler_1962}, $\varepsilon_{\mathrm{SL}}$ is the
$\mathbf{q} \to 0$ and $\omega \to 0$ limits of the non-interacting
dielectric function $\varepsilon(\mathbf{q}, \omega)$, where
$\mathbf{q}$ is the momentum transfer and $\omega$ is the frequency:
\begin{equation}
  \label{eq:RPA-eps2}
  \varepsilon_{\mathrm{SL}}
  = \lim_{\mathbf{q} \to 0} 1 + \frac{2e^{2}}{\varepsilon_{0} |\mathbf{q}|^{2} \Omega}
  \sum_{\mathrm{k, c, v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
  \left[f(\psi_{\mathrm{c}}) - f(\psi_{\mathrm{v}})\right]
\end{equation}
where $e$ is the unit charge, c, v are the conduction and valence
bands, $E$ is the eigenenergy of individual bands, and $f$ is the
Fermi-Dirac distribution function. Taking the
limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
at 0 K can be unified by the same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$. Next we will show
that the different behavior of $\psi^{\parallel}$ and $\psi^{\perp}$
give rise to the main text Eqs. \ref{main-eq:2D-Moss-para} and
\ref{main-eq:2D-Moss-perp}.

\subsubsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-para}}
\label{sssec:theory-1}

In this section we show how main text Eq. \ref{main-eq:2D-Moss-para}
is derived from Eq. \ref{eq:alpha-RPA}. For the in-plane component
$\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$ is independent
of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes the independent of $\psi^{\perp}$, due to the orthogonality
and normality. The Bloch-wave form of $\psi^{\parallel}$ ensures that
only the cell-function $u(\mathbf{k})$ contributes to the final result
of $\alpha^{\parallel}_{\mathrm{2D}}$ \cite{davies_physics_1997}, such
that:
\begin{equation}
  \label{eq:alpha_para_RPA}
  \alpha_{\mathrm{2D}}^{\parallel} = \frac{2e^{2}}
  {(2 \pi)^{2}} \int d^{2}\mathbf{k} \sum_{\mathrm{c, v}}
  \frac{|<u_{\mathrm{v}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
Present the summation of k-space into integral, and further combining
with the Bloch wave representation and the
$\mathbf{k} \cdot \mathbf{p}$ theory\cite{Jiang_2017_Eg_Eb}, we arrive
at Eq. \ref{main-eq:2D-Moss-para}. Note that from
Ref. \citenum{Jiang_2017_Eg_Eb}, the coefficient between
$\alpha^{\parallel}_{\mathrm{2D}}$ and $E_{\mathrm{g}}$ is given by:
$\alpha_{\mathrm{2D}}^{\parallel} = N e^{2}/(2\pi E_{\mathrm{g}})$,
where $N$ is the degeneracy associated with the minimal bandgap. Such
relation predicts the linear correlation coefficient between
$(4\pi \varepsilon_{0})/\alpha^{\parallel}_{\mathrm{2D}}$ (measured in
\AA{}$^{-1}$) and $E_{\mathrm{g}}$ (measured in eV) as
$8 \pi^{2} \varepsilon_{0} \AA{} / (e N) \approx 0.436 / N$. From main
text Figure \ref{main-fig-3}c, we have
$(4 \pi \varepsilon_{0}\cdot \AA{})/\alpha_{\mathrm{2D}}^{\parallel}
\sim{} 0.183 E_{\mathrm{g}}/ \mathrm{eV}$, corresponding to $N$
between 2 and 3, which is a reasonable result for the 2D materials
studied.

\subsubsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-perp}}
\label{ssec:theory-2}
For main text Eq. \ref{main-eq:2D-Moss-perp}, treating the in-plane
wavefunctions as plane wave with form
$\psi^{\parallel}(\rho) \propto e^{i \mathbf{k \rho}}$, the matrix
element of
$<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>$,
when $\mathbf{q}=(0, 0, q_{z})$, becomes\cite{Hybertsen_1987}:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
  <\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>
  &= \frac{1}{A} \int dx \int dy
  e^{i(\mathbf{-k \rho} - \mathbf{q \rho} + \mathbf{(k+q) \rho})}
  \int (\psi^{\perp})^{*}_{\mathrm{v}}(\mathbf{k})e^{-iq_{z}z}\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})\\
  &= <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})>
  \end{aligned}
\end{equation}
Note that the states perpendicular are bound, the integral is
meaningful only when $\mathbf{k=k+q}$ \cite{davies_physics_1997}. By
performing the Taylor expansion of
$e^{-i\mathbf{qr}} \approx 1 - i\mathbf{qr}$, we get:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z-2}
    <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
    &\approx <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})> -
    iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>\\
    &= -iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
   \end{aligned}
\end{equation}
plug this into Eq. \ref{eq:alpha-RPA} and express the summation over
$k_{x}$ and $k_{y}$ in a continuous form within the Brillouin Zone, we
arrive at:
\begin{equation}
\label{eq:alpha_perp_RPA}
\alpha_{\mathrm{2D}}^{\perp} = \frac{2e^{2}}{(2 \pi) ^{2}} \int d^{2}\mathbf{k}
\sum_{\mathrm{c, v}}
\frac{|<\psi_{\mathrm{v}}(\mathbf{k})|z|\psi_{\mathrm{c}}(\mathbf{k})>|^{2}}
{E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
The formalism is slightly different from Eq.\ref{eq:alpha_para_RPA}.

The out-of-plane wavefunction $\psi^{\perp}(z)$ is the solution to the Schr\"{o}dinger equation with
Hamiltonian $\mathcal{H} = -\hbar^{2} \nabla^{2}/2m_{e} + V(z)$, where
$\hbar$ is the reduced Planck constant, $m_{e}$ is electron mass and
$V(z)$ is the confined Coulombic potential along the z-direction
created by the nuclei
\cite{davies_physics_1997,ihn_semiconductor_2009}. Although the exact
form for $\psi^{\perp}$ depends on the exact distribution of $V(z)$,
without loss of generality we can assume the electrons are confined in
a potential well of width $\delta$, which is the typical treatment for
semiconductor QWs \cite{ihn_semiconductor_2009}. The allowed bound
states inside the confined region generally has wave vector
$k_{z} \propto n \pi / \delta$. With the total energy
$E_{n}(\mathbf{k}) = {\displaystyle \frac{\hbar^{2} (k_{x}^{2} +
    k_{y}^{2})}{2 m^{\parallel}} + \frac{\hbar^{2} n^{2} \pi^{2}}{2
    m^{\perp} \delta^{2}}}$, where $m^{\parallel}$ and $m^{\perp}$ are
the effective masses parallel and perpendicular to the 2D
plane. Therefore the denominator of Eq. 10 becomes independent of
$\mathbf{k}$, that
$E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k}) =
(n_{\mathrm{c}}^{2} - n_{\mathrm{v}}^{2}) {\displaystyle
  \frac{\hbar^{2} \pi^{2}}{2 m^{\perp} \delta^{2}}}$. On the other
hand, the numerator
$<\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>$
is proportional to the confinement length $\delta$ (can be seen using
particle-in-box solution\cite{davies_physics_1997}). In combination,
the individual terms of the summation in the right hand of Eq. 5b is
independent of neither $E_{\mathrm{g}}$ nor \textbf{k}, proving that
$\alpha_{\mathrm{2D}}^{\perp}$ is independent of the band gap.

Finally we will show the $\alpha_{\mathrm{2D}}^{\perp}$ is related to the
confinement length $\delta$, or equivalent the thickness of the 2D
material. For simplicity we assume $\psi^{\parallel}$ is the solution
of a particle confined inside an infinite potential well of width
$\delta$ with the form $\psi \propto \sin(\frac{n\pi}{\delta})$. At 0
K, the only non-zero contribution for the transition between band
$n_{\mathrm{v}}$ and $n_{\mathrm{c}}$ corresponds with in-plane wave
vector $\mathbf{k}_{\rho}$ within the range of $k_{\mathrm{F}}^{\mathrm{v}}$ to
$k_{\mathrm{F}}^{\mathrm{c}}$, the Fermi vectors of each band. Therefore we have:
\begin{equation}
  \begin{aligned}
    \alpha_{\mathrm{2D}}^{\perp} &\propto \delta^{4} m^{\perp} \int
    d^{2}\mathbf{k}_{\rho}
    \frac{f(\psi_{\mathrm{v}}(\mathbf{k}_{\rho}))
      -f(\psi_{\mathrm{v}}(\mathbf{k}_{\rho})) }{n_{\mathrm{c}}^{2} -
      n_{\mathrm{v}}^{2}}\\
    &= \delta^{4} m^{\perp} \sum_{\mathrm{c, v}}\pi \frac{
      (k_{\mathrm{F}}^{\mathrm{c}})^{2} -
        (k_{\mathrm{F}}^{\mathrm{v}})^{2}}{n_{\mathrm{c}}^{2} -
        n_{\mathrm{v}}^{2}}
  \end{aligned}
\end{equation}
Since the Fermi wave vector for band \textit{i} follows
${\displaystyle
  \frac{\hbar^{2}(k_{\mathrm{F}}^{i})^{2}}{2m^{\parallel}}} +
{\displaystyle \frac{\hbar^{2} n_{i}^{2} \pi^{2}}{2m^{\perp}
    \delta^{2}}} = E_{\mathrm{F}}$, we have
$ (k_{\mathrm{F}}^{\mathrm{c}})^{2} -
(k_{\mathrm{F}}^{\mathrm{v}})^{2} \propto (n_{\mathrm{c}}^{2} -
n_{\mathrm{v}}^{2}) {\displaystyle
  \frac{m^{\parallel}}{m^{\perp}}}$. Note that in the particle-in-box
model, the energy levels for a sub-nm QW will be at the order of 10
eV, which allows only few states inside the QW. Combine with the fact
that the strongest transition mostly comes from the lowest bands
\cite{davies_physics_1997}, $\alpha^{\perp}_{\mathrm{2D}}$ is shown to
be proportional to $\delta^{p}$, where the power index $p$ is 2. We
note the different power law comes from the choice of the $V(z)$. The
atomic polarizabilities of free particles inside a 1D infinite QW,
finite QW or harmonic QW all show dependency with $\delta^{4}$, where
$\delta$ is the length of confinement or effective harmonic length of
the QW \cite{Fowler_1984,Maize_2011}. Since the 2D polarizability is
the molecular polarizability divided by the area, such approach will
always yield $\alpha_{\mathrm{2D}}^{\perp} \propto \delta^{2}$. To
accurately model the $\alpha_{\mathrm{2D}}^{\perp}$ we need a better
model for the confinement in z-direction than simple square potential
well, which may be tedious. Nevertheless the simple model presented
here still captures the bandgap-indepent and thickness-related feature
of $\alpha_{\mathrm{2D}}^{\perp}$, supporting our findings in the main
text. In the next section we will provide a simple explanation for the $\alpha \propto \delta$ relation from fundamental electrostatics theory.

\subsubsection{Explanation to main text Eq. \ref{main-eq:2D-Moss-perp}
  from fundamental electrostatics}
\label{sssec:theory-3}

The dependency of $\alpha_{\mathrm{2D}}^{\perp}$ on the thickness
$\delta$ of a 2D material, can also be regarded using fundamental
electrostatic model. Consider the smallest repeating unit of the 2D
material with xy-plane area $A$, under small perturbation field $E$
along the z-direction.  Note that the surface bound charge
$\sigma_{\mathrm{b}}=n e /A$, where $n$ is the number of unit charges
contributes to the bound charges, comes only from the dipoles of the
outer-most atoms, since the induced charges from inner atoms are
cancel out (see Figure \ref{fig:classic-model}).
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7\linewidth]{img/SI-classic.pdf}
  \caption{Fundamental electrostatic model for the
    thickness-dependency of $\alpha_{\mathrm{2D}}^{\perp}$, using 2H-MoS2 as an
    example. Left: induced dipoles from individual atoms along the
    z-direction. The positive and negative induced charges from inner
    atoms cancel out. Right: simplied model for the thickness
    dependency of $\alpha_{\mathrm{2D}}^{\perp}$, where the surface dipole density
    $\boldsymbol{\mu}$ comes only from the outer-most atoms.}
  \label{fig:classic-model}
\end{figure}
From the definition of
$\alpha_{\mathrm{2D}}^{\perp}$, we have:
\begin{equation}
  \label{eq:alpha-classic}
  \alpha_{\mathrm{2D}}^{\perp} = \frac{\boldsymbol{u}_{z}}{E_{\mathrm{loc}} A}
  = \frac{(d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j}) \sigma_{\mathrm{b}}}{E}
\end{equation}
where $r_{\mathrm{cov}}^{i}$ and $r_{\mathrm{cov}}^{j}$ are the
covalent radii of the outer-most atoms, the characteristic length of
the dipole extension in z-direction, and $d_{\mathrm{max}}$ is the
z-distance between the nuclei of such atoms.  The field $E$
counterbalances the field from the surface bound charges and equals
$E = \sigma_{\mathrm{b}}/\varepsilon_{0}$. Therefore we have:
\begin{equation}
  \label{eq:alpha-classic-2}
  \alpha_{\mathrm{2D}}^{\perp} = (d_{\mathrm{max}} + r_{\mathrm{cov}}^{i} + r_{\mathrm{cov}}^{j})\varepsilon_{0}
                = \delta_{\mathrm{cov}} \varepsilon_{0}
\end{equation}
which explains the linear relation seen in main text Figure
\ref{main-fig-3}d. We can see that such simple model nicely captures
the thickness feature of $\alpha_{\mathrm{2D}}^{\perp}$, and
reproduces the right coefficient between $\delta_{\mathrm{cov}}$ and
$\alpha_{\mathrm{2D}}^{\perp}$. Finally, the schematic drawing of the
polarizability ellipsoid for a 2D material, is drawn in Figure
\ref{fig:ellipsoid}.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{img/SI-ellipsoid.pdf}
  \caption{Polarizability ellipsoid of a 2D material, the long axis
    equals $\alpha_{\mathrm{2D}}^{\parallel}/2\varepsilon_{0}$ and the short axis
    equals  $\alpha_{\mathrm{2D}}^{\perp}/2\varepsilon_{0}$}
  \label{fig:ellipsoid}
\end{figure}

\subsection{}
\label{sec:theory-2}
